The Unicode Cookbook \linebreak for Linguists

### 2018-03-07 book/chapters/implementation.tex
all but the final reviewer comments on the specification

\section{Overview}
To illustrate the practical applications of orthography profiles, as defined in Chapter~\ref{orthography-profiles}, we have implemented two versions of the specifications presented there: one in Python\footnote{\url{https://pypi.python.org/pypi/segments}} and one in R.\footnote{\url{https://github.com/cysouw/qlcData}} This chapter introduces these two implementations, and gives practical step-by-step guidelines for installing and using them. Various simple and sometimes somewhat abstract examples will be discussed to show the different options available, and to illustrate the intended usage of orthography profiles.

Note that our two software libraries have rather different implementation histories, thus they may not give the same results in all situations. We do, however, provide a test suite for each implementation aimed at making sure that the results covered by those test cases are identical. Users should refer to the tests and to the documentation in each release for specifics regarding each implementation. Due to the different naming conventions in Python and R, function names differ across the two implementations. The performance with larger datasets also differs per implementation. In sum, the two software libraries should be considered as proof of concept and not as the final word on the practical application of the specifications above. In our own experience, the current versions are sufficiently fast and stable to be useful for academic practice (e.g.\ checking data consistency, or analyzing and transliterating small to medium sized data sets), but they should probably not be used for full-scale industry applications without some adaptations.

In Section \ref{installing-python-and-r}, we first explain how to install Python\footnote{\url{https://www.python.org/}} and R.\footnote{\url{https://www.r-project.org/}} In Sections \ref{python-implementations} and \ref{r-implementation}, we discuss our Python and R software packages, respectively. In addition to the material presented here to get users started, we maintain several case studies that use orthography profiles in action. For convenience, we make these use cases available as Jupyter Notebooks\footnote{\url{http://jupyter.org/}} in our Github repository of code recipes, which accompanies this book online.\footnote{\url{https://github.com/unicode-cookbook/}} In the final section in this chapter, we briefly describe these additional use-cases and point the user to the accompanying code online.

@@

You can also install the \texttt{segments} package with from the Github repository:

@@

\subsection*{Application programming interface}
The \texttt{segments} API can be accessed by importing the package into Python or by writing a Python script. Here w\is an example of how to import the libraries, create a tokenizer object, tokenize a string, and create an orthography profile.

\begin{lstlisting}[language=python]
>>> from segments.tokenizer import Tokenizer
\end{lstlisting}

@@

% \begin{lstlisting}[language=python]
\begin{lstlisting}[inputencoding=utf8,extendedchars=true,escapeinside={\%*}{*)},texcl=true]
>>> t = Tokenizer()
>>> result = t.characters("%*ĉháɾã̌ctʼɛ↗ʐː| k͡p*")
>>> print(result)
>>> %*c ̂ h a ́ ɾ a ̃ ̌ c t ʼ ɛ ↗ ʐ ː | # k ͡ p
\end{lstlisting}

\begin{lstlisting}[inputencoding=utf8,extendedchars=true
  basicstyle=\footnotesize\ttfamily,
  tabsize=2,
  breaklines=true,
  language=python,
  keywordstyle=\color{blue!60!black},
  stringstyle=\color{green!60!black}]
>>> t = Tokenizer()
>>> result = t.characters("ĉháɾã̌ctʼɛ↗ʐː| k͡p")
>>> print(result)
>>> c ̂ h a ́ ɾ a ̃ ̌ c t ʼ ɛ ↗ ʐ ː | # k ͡ p
\end{lstlisting}

@@

\begin{lstlisting}[language=python,texcl=true]
>>> result = t.grapheme_clusters("ĉháɾã̌ctʼɛ↗ʐː| k͡p")
>>> print(result)
>>> ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p
\end{lstlisting}

\noindent The \texttt{grapheme\_clusters} function is also the default segmentation algorithm for the \texttt{segments.Tokenizer}. It is useful when you encounter a text that you want to tokenize to identify orthographic or transcription elements.

\begin{lstlisting}[language=python]
>>> result = t("ĉháɾã̌ctʼɛ↗ʐː| k͡p")
>>> print(result)
>>> ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p
\end{lstlisting}

\noindent The \texttt{ipa} parameter forces grapheme segmentation for IPA strings (a formal definition is given in Chapter \ref{orthography-profiles}.

\begin{lstlisting}[language=python]
>>> result = t("ĉháɾã̌ctʼɛ↗ʐː| k͡p", ipa=True)
>>> print(result)
>>> ĉ h á ɾ ã̌ c tʼ ɛ ↗ ʐː | # k͡p
\end{lstlisting}

\noindent We can also load an orthography profile and tokenize input string with it. In the data directory we've placed an example orthography profile. Let's have a look at it using \texttt{more} on the command line.

\begin{lstlisting}[language=bash,texcl=true]
  $ more data/orthography-profile.tsv
  Grapheme  IPA   XSAMPA  COMMENT
  a         a     a
  aa        aː    a:
  b         b     b
  c         c     c
  ch        tʃ    tS
  -         NULL  NULL    "comment with   tab"
  on        õ     o~
  n         n     n
  ih        í     i_H
  inh       ĩ́     i~_H
\end{lstlisting}

@@

\begin{lstlisting}[language=python]
>>> from segments.tokenizer import Profile
>>> t = Tokenizer('data/orthography-profile.tsv')
\end{lstlisting}

\noindent Now let's segment the graphemes in some input strings with our orthography profile. The output is segmented given the definition of graphemes in our orthography profile, e.g.\ we specified the sequence of two $<$a a$>$ should be a single unit $<$aa$>$, and so should the sequences $<$c h$>$, $<$o n$>$ and $<$i h$>$.

\begin{lstlisting}[language=python]
>>> t('aabchonn-ih')
>>> 'aa b ch on n - ih'
\end{lstlisting}

@@

\begin{lstlisting}[language=python]
>>> t.transform("aabchonn-ih", "IPA")
>>> 'aː b tʃ õ n í'
\end{lstlisting}

\begin{lstlisting}[language=python]
>>> t.transform("aabchonn-ih", "XSAMPA")
>>> 'a: b tS o~ n i_H'
\end{lstlisting}

@@

\begin{lstlisting}[language=python]
>>> t = Tokenizer('data/orthography-profile.tsv', 
	errors_replace=lambda c: '?')
>>> t.find_missing_characters("aa b ch on n - ih x y z")
>>> 'aa b ch on n - ih ? ? ?'
\end{lstlisting}

\begin{lstlisting}[language=python]
>>> t = Tokenizer('data/orthography-profile.tsv', 
	errors_replace=lambda c: '<{0}>'.format(c))
>>> t.find_missing_characters("aa b ch on n - ih x y z")
>>> 'aa b ch on n - ih <x> <y> <z>'
\end{lstlisting}

\noindent Perhaps you want to create an initial orthography profile that also contains those graphemes x, y, z?

\begin{lstlisting}[language=python]
>>> profile = Profile.from_text("aa b ch on n - ih x y z")
>>> print(profile)
\end{lstlisting}

@@

\begin{comment}
\begin{lstlisting}[language=python]
>>> from __future__ import unicode_literals, print_function
>>> from segments import Profile, Tokenizer
>>> t = Tokenizer()
>>> t('abcd')
'a b c d'
>>> prf = Profile({'Grapheme': 'ab', 'mapping': 'x'}, \
{'Grapheme': 'cd', 'mapping': 'y'})
>>> print(prf)
mapping Grapheme
ab  x
cd  y
>>> t = Tokenizer(profile=prf)
>>> t('abcd')
'ab cd'
>>> t('abcd', column='mapping')
'x y'
\end{comment}

@@

\begin{lstlisting}[language=bash]
  $ segments -h
\end{lstlisting}

@@

  $ cat text.txt | segments --mapping=mapping 
    --profile=profile.prf tokenize
  a ä x ö x ü x
\end{lstlisting}

@@

<<size='scriptsize', engine='bash', tidy=FALSE>>=
tokenize --help
@

@@

Note that these different profiles can also be produced using the bash
executable \texttt{writeprofile} (see \ref{r-implemenation} for notes
on how to install the bash executable). Exactly this example is also included in
the help file of the executable.

<<size='scriptsize', engine='bash'>>=
writeprofile --help
@

@@

<<eval=FALSE, tidy=FALSE>>=
# add a contextual grapheme, and then use the edited profile to tokenize
tokenize( example
         , profile = "~/Desktop/profile_skeleton.txt"
         , regex = TRUE
         , ordering = "context"
        )$strings
@

@@

<<eval=FALSE, tidy=FALSE>>=
# add a contextual grapheme, and then use the edited profile to tokenize
tokenize( example
         , profile = "~/Desktop/profile_skeleton.txt"
         , regex = TRUE
         , transliterate = "IPA"
        )$strings
@

@@

\section{More recipes}
\label{use-cases}

Online we provide several recipes, or use cases, that illustrate the applications of orthography profiles using our implementations in Python and R. \footnote{\url{https://github.com/unicode-cookbook/recipes}} Here we briefly describe these use cases and we encourage users to try them out using Git and Jupyter Notebooks.

First, as we discussed above, we provide a basic tutorial on how to use the Python \texttt{segments}\footnote{\url{https://pypi.python.org/pypi/segments}} and R \texttt{qlcData}\footnote{\url{https://github.com/cysouw/qlcData}} libraries.\footnote{\url{https://github.com/unicode-cookbook/recipes/tree/master/Basics}} This use case simply shows the basic functions of each library.

The two use cases using the Python \textt{segments} package include a tutorial on how to segment graphemes in IPA text:

@@

\noindent The JIPA use case uses excerpts from the Northwind and the Sun passages from the Illustrations of the IPA published in the Journal of the International Phonetic Alphabet. Thus the use case shows how a user might tokenize IPA proper. The Dogon use case uses fieldwork data from the Dogon languages of Mali language documentation project.\footnote{\url{http://dogonlanguages.org/}} This use case illustrates how a user might tokenize fieldwork data from numerous linguists using different transcription practices by defining these practices with an orthography profile to make the output unified and comparable.

The two use cases using the R \texttt{qlcData} library include a use case for tokenizing wordlist data from the Automated Similarity Judgment Program (ASJP):\footnote{\url{http://asjp.clld.org/}}

@@

\noindent The ASJP use case shows how to download the full set of ASJP wordlists, to combine them into a single large CSV file, and to tokenize the ASJP orthography. The Dutch use case uses the 10K corpus for Dutch (``nld'') from the Leipzig Corpora Collection,\footnote{\url{http://corpora2.informatik.uni-leipzig.de/download.html}} which is then cleaned and tokenized with an orthography profile that captures the intricacies of Dutch orthography.

>>>>

\section{Overview}
To illustrate the practical applications of orthography profiles, we have implemented two versions of the specifications presented in Chapter~\ref{orthography-profiles}: one in Python\footnote{\url{https://pypi.python.org/pypi/segments}} and one in R.\footnote{\url{https://github.com/cysouw/qlcData}} In this chapter, we introduce these two software libraries and provide practical step-by-step guidelines for installing and using them. Various simple and sometimes somewhat abstract examples will be discussed to show the different options available, and to illustrate the intended usage of orthography profiles in general.

Note that our two libraries have rather different implementation histories, thus they may not give the same results in all situations (as discussed in Chapter~\ref{orthography-profiles}). However, we do provide extensive test suites for each implementation that follow standard practices to make sure that results are correct. Users should refer to these tests and to the documentation in each release for specifics about each implementation. Note that due to the different naming convention practices in Python and R, function names differ between the two libraries. Also, the performance with larger datasets may not be comparable between the Python and R implementations. In sum, our two libraries should be considered as proof of concept and not as the final word on the practical application of the specifications discussed in the previous chapter. In our experience, the current versions are sufficiently fast and stable to be useful for academic practice (e.g.\ checking data consistency, or analyzing and transliterating small to medium sized data sets), but they should probably not be used for full-scale industry applications without adaptation.

First, in Section \ref{installing-python-and-r} we explain how to install Python\footnote{\url{https://www.python.org/}} and R.\footnote{\url{https://www.r-project.org/}} Then in Sections \ref{python-implementations} and \ref{r-implementation}, we discuss our Python and R software packages, respectively. In addition to the material presented here to get users started, we maintain several case studies online that illustrate how to use orthography profiles in action. For convenience, we make these recipes available as Jupyter Notebooks\footnote{\url{http://jupyter.org/}} in our Github repository.\footnote{\url{https://github.com/unicode-cookbook/}} In the final section in this chapter, we also briefly describe a few recipes that we do not go into detail in this book.

@@

You can also install the \texttt{segments} package with from the Github repository:\footnote{\url{https://github.com/cldf/segments}}

@@

\subsection*{Application programming interface}
The \texttt{segments} API can be accessed by importing the package into Python or by writing a Python script. Here is an example of how to import the libraries, create a tokenizer object, tokenize a string, and create an orthography profile.

\begin{lstlisting}
>>> from segments.tokenizer import Tokenizer
\end{lstlisting}

@@

% \lstset{extendedchars=false, escapeinside=**}
\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t = Tokenizer()
>>> result = t.characters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@c ̂ h a ́ ɾ a ̃ ̌ c t ʼ ɛ ↗ ʐ ː | # k ͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t.grapheme_clusters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

\noindent The \texttt{grapheme\_clusters} function is the default segmentation algorithm for the \texttt{segments.Tokenizer}. It is useful when you encounter a text that you want to tokenize to identify orthographic or transcription elements.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

\noindent The \texttt{ipa} parameter forces grapheme segmentation for IPA strings.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', ipa=True)
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c tʼ ɛ ↗ ʐː | # k͡p@*)'
\end{lstlisting}

\noindent We can also load an orthography profile and tokenize an input string with it. In the data directory we've placed an example orthography profile. Let's have a look at it using \texttt{more} on the command line.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
  $ more data/orthography-profile.tsv
  Grapheme  IPA   XSAMPA  COMMENT
  a         a     a
  aa        (*@aː@*)    (*@a:@*)
  b         b     b
  c         c     c
  ch        (*@tʃ@*)    tS
  -         NULL  NULL    "comment with   tab"
  on        (*@õ@*)     o~
  n         n     n
  ih        (*@í@*)     i_H
  inh       (*@ĩ́@*)     i~_H
\end{lstlisting}

@@

\begin{lstlisting}
>>> from segments.tokenizer import Profile
>>> t = Tokenizer('data/orthography-profile.tsv')
\end{lstlisting}

\noindent Now let's segment the graphemes in some input strings with our orthography profile. The output is segmented given the definition of graphemes in our orthography profile, e.g.\ we specified the sequence of two <a a> should be a single unit <aa>, and so should the sequences <c h>, <o n> and <i h>.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t('(*@aabchonn-ih@*)')
>>> '(*@aa b ch on n - ih@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t.transform('(*@aabchonn-ih@*)', 'IPA')
>>> '(*@aː b tʃ õ n í@*)'
\end{lstlisting}

% For some reason that fails me, this does not work:
% \begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
% >>> t.transform('aabchonn-ih', 'XSAMPA')
% >>> '(*@a: b tS o~ n i_H@*)'
% \end{lstlisting}

\begin{lstlisting}
>>> t.transform('aabchonn-ih', 'XSAMPA')
>>> 'a: b tS o~ n i_H'
\end{lstlisting}

@@

\begin{lstlisting}
>>> t = Tokenizer('data/orthography-profile.tsv', 
	errors_replace=lambda c: '?')
>>> t.find_missing_characters("aa b ch on n - ih x y z")
>>> 'aa b ch on n - ih ? ? ?'
\end{lstlisting}

\begin{lstlisting}
>>> t = Tokenizer('data/orthography-profile.tsv', 
	errors_replace=lambda c: '<{0}>'.format(c))
>>> t.find_missing_characters("aa b ch on n - ih x y z")
>>> 'aa b ch on n - ih <x> <y> <z>'
\end{lstlisting}

\noindent Perhaps you want to create an initial orthography profile that also contains those graphemes x, y, z? Note that the space character and its frequency are also captured in this initial profile.

\begin{lstlisting}
>>> profile = Profile.from_text("aa b ch on n - ih x y z")
>>> print(profile)
\end{lstlisting}

@@

@@

\begin{lstlisting}[language=bash]
  $ segments -h

usage: segments [-h] [--verbosity VERBOSITY] [--encoding ENCODING]
                [--profile PROFILE] [--mapping MAPPING]
                command ...

Main command line interface of the segments package.

positional arguments:
  command               tokenize | profile
  args

optional arguments:
  -h, --help            show this help message and exit
  --verbosity VERBOSITY
                        increase output verbosity
  --encoding ENCODING   input encoding
  --profile PROFILE     path to an orthography profile
  --mapping MAPPING     column name in ortho profile to map 
                        graphemes

Use 'segments help <cmd>' to get help about individual commands.  
\end{lstlisting}

\noindent We have created some test data\footnote{\url{https://github.com/unicode-cookbook/recipes/tree/master/Basics/sources}} with the German word \textit{Schächtelchen}, which is the diminuitive form of \textit{Schachtel}, meaning `box', `packet', or `carton' in English.

\begin{lstlisting}[language=bash]
  $ more sources/german.txt

  Schächtelchen
\end{lstlisting}

\noindent We can create an initial orthography profile of the German text by passing it to the \textsc{segments profile} command. The initial profile tokenizes the text on Unicode grapheme clusters, lists the frequency of each grapheme, and provides an initial mapping column by default.

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments profile

  Grapheme  frequency  mapping
  c         3          c
  h         3          h
  e         2          e
  S         1          S
  (*@ä@*)         1          (*@ä@*)
  t         1          t
  l         1          l
  n         1          n
\end{lstlisting}

\noindent Next, we know a bit about German orthography and which characters combine to form German graphemes. We can use the information from our initial orthography profile to hand-curate a more precise German orthography profile that takes into account capitalization (German orthography obligatorily capitalizes nouns) and grapheme clusters, such as <sch> and <ch>. We can use the initial orthography profile above as a starting point (note the in large text the frequency column may signal errors in the input, such as typos, if they occur with a very low frequency). The initial orthography profile can be edited with a text editor or spreadsheet program. As per the orthography profile specifications (see Chapter \ref{orthography-profiles}), we can adjust rows in the \textsc{Grapheme} column and then add additional columns for translitations or comments.

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ more more data/german-orthography-profile.tsv

  Grapheme  IPA  XSAMPA  COMMENT                     
  Sch       (*@ʃ@*)    S       German nouns are capitalized
  (*@ä@*)         (*@ɛː@*)   E:                                  
  ch        (*@ç@*)    C                                   
  t         t    t                                   
  e         e    e                                   
  l         l    l                                   
  n         n    n                                    
\end{lstlisting}

\noindent Using the command line \textsc{segments} function and passing it our orthography profile, we can now segment our German text example into graphemes.

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments 
    --profile=data/german-orthography-profile.tsv tokenize

  '(*@Sch ä ch t e l ch e n@*)'
\end{lstlisting}

\noindent By providing \textsc{segments} a column for transliteration, we can convert the text into IPA.

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments --mapping=IPA 
	--profile=data/german-orthography-profile.tsv tokenize

  '(*@ʃ ɛː ç t e l ç e n@*)'
\end{lstlisting}

\noindent And we can transliterate to XSAMPA.

\begin{lstlisting}[language=bash]
  $ cat sources/german.txt | segments --mapping=XSAMPA 
	--profile=data/german-orthography-profile.tsv tokenize

  'S E: C t e l C e n'
\end{lstlisting}

\noindent More examples are available online.\footnote{\url{https://github.com/unicode-cookbook/recipes}}

% Old example that reviewer didn't like!
\begin{comment}
\begin{lstlisting}[language=bash]
  $ more text.txt
  aäaaöaaüaa
\end{lstlisting}

@@

  $ cat text.txt | segments --mapping=mapping 
    --profile=profile.prf tokenize
  a ä x ö x ü x
\end{lstlisting}
\end{comment}
% \end old example that reviewer didn't like!

@@

% TODO:
% <<size='scriptsize', engine='bash', tidy=FALSE>>=
% tokenize --help
% @

@@

Note that these different profiles can also be produced using the bash
executable \texttt{writeprofile} (see above for how to install the bash executable). 
Exactly this example is also included in the help file of the executable.

% TODO: fix
% <<size='scriptsize', engine='bash'>>=
% writeprofile --help
% @

@@

<<eval=FALSE, tidy=FALSE>>=
# add a contextual grapheme, and then use the edited 
# profile to tokenize
tokenize( example
         , profile = "~/Desktop/profile_skeleton.txt"
         , regex = TRUE
         , ordering = "context"
        )$strings
@

@@

<<eval=FALSE, tidy=FALSE>>=
# add a contextual grapheme, and then use the edited 
# profile to tokenize
tokenize( example
         , profile = "~/Desktop/profile_skeleton.txt"
         , regex = TRUE
         , transliterate = "IPA"
        )$strings
@

@@

\section{Recipes online}
\label{use-cases}

Online we provide several use cases -- what we refer to as recipes -- that illustrate the applications of orthography profiles using our implementations in Python and R.\footnote{\url{https://github.com/unicode-cookbook/recipes}} Here we briefly describe these use cases and we encourage users to try them out using Git and Jupyter Notebooks.

First, as we discussed above, we provide a basic tutorial on how to use the Python \texttt{segments}\footnote{\url{https://pypi.python.org/pypi/segments}} and R \texttt{qlcData}\footnote{\url{https://github.com/cysouw/qlcData}} libraries. This recipe simply shows the basic functions of each library to get you started.\footnote{\url{https://github.com/unicode-cookbook/recipes/tree/master/Basics}}

The two recipes using the Python \textt{segments} package include a tutorial on how to segment graphemes in IPA text:

@@

\noindent The JIPA recipes uses excerpts from the Northwind and the Sun passages from the Illustrations of the IPA published in the Journal of the International Phonetic Alphabet. Thus the recipe shows how a user might tokenize IPA proper. The Dogon recipe uses fieldwork data from the Dogon languages of Mali language documentation project.\footnote{\url{http://dogonlanguages.org/}} This recipe illustrates how a user might tokenize fieldwork data from numerous linguists using different transcription practices by defining these practices with an orthography profile to make the output unified and comparable.

The two recipes using the R \texttt{qlcData} library include a use case for tokenizing wordlist data from the Automated Similarity Judgment Program (ASJP):\footnote{\url{http://asjp.clld.org/}}

@@

\noindent The ASJP use case shows how to download the full set of ASJP wordlists, to combine them into a single large CSV file, and to tokenize the ASJP orthography. The Dutch use case takes as input the 10K corpus for Dutch (``nld'') from the Leipzig Corpora Collection,\footnote{\url{http://corpora2.informatik.uni-leipzig.de/download.html}} which is then cleaned and tokenized with an orthography profile that captures the intricacies of Dutch orthography.

% In closing, using Github to share code and data provides a platform for sharing scientific results and it also promotes a means for scientific replicability of results. Moreover, we find that in cases where the scientists are building the to tools for analysis, open repositories and data help to ensure that what you see is what you get.
